---
title: "DebrisPerSite"
author: "Brie Sherow"
date: "15/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r figure-setup, echo=FALSE, include=FALSE} 
knitr::opts_chunk$set(fig.path="Output/", fig.width=6.5,fig.height=4, fig.align="center")
```

```{r load libraries, echo=FALSE, include=FALSE}
lib.load = c("here", "ggplot2", "tidyverse", "lubridate") 
lapply(lib.load, require, character.only = TRUE) 
```

```{r load data}
#set working directory
setwd("~/Desktop/Marine Debris/R-project/marine-debris")
getwd()

#Check working directory
here::here()

#Load data, assign date to date and notes to character
data <- read.csv(file="data/all_data.csv", header=T, sep=",")
data$Date <- dmy(data$Date)
data$Notes <- as.character(data$Notes)

glimpse(data)
```


```{r filter sites}
#filtering for sites surveyed at least 3 times and by UNSW teams
site_select <- data %>%
  filter(Team=="UNSW" & Location %in% c("DOUBL","WATSO","PARSL","NEUTR","CLIFT","ATHOL")) %>%
  select(-Habitat2, -Fragment, -Whole)
```

```{r normalise survey count}
#create a count for number of surveys
sampling_dates <- data %>%
  group_by(Location) %>%
  filter(Team=="UNSW") %>%
  count(Date)

#take away the debris count
sampling_dates$n <- NULL

sampling_visits <- sampling_dates %>%
  count(Location) %>%
  arrange(desc(n)) %>%
  rename(NumberSurveys=n)

#join survey to site select
site_select <- left_join(site_select, sampling_visits, by="Location")

#add unique event id
site_select <- site_select %>%
  mutate(Event_ID=paste(Date, Site))

write.csv(site_select, file="Output/siteselect.csv")
```

```{r plot setup}
#Total item count of all IDs per survey event
event_total <- site_select %>%
  group_by(Event_ID) %>%
  mutate(event_sum=sum(Total), #total debris count per event
         event_prop=Total/event_sum)%>%  #total number of all items per survey event
  ungroup()

#normalise sites
site_total <- event_total %>%
  group_by(Site, ID) %>%
  mutate(mean_prop=sum(event_prop)/NumberSurveys, #mean of item type per site
         proportion_sd=sd(event_prop)) %>% #sd of item type per site
        ungroup()

#remove unnecessary data for plotting
site_plots <-
  site_total %>%
  filter(mean_prop>0) %>%
  select(Site, Location, Habitat, ID, Material, mean_prop, proportion_sd) %>%
  distinct()
```


```{r top item}
#total item count per ID
item_count <- site_plots %>%

```{r top item by pier}
#total item count per ID
item_count_p <- site_plots %>%
  filter(Habitat=="P") %>%
  count(ID) %>%
  arrange(desc(n)) %>%
  rename(IDCount=n)

#join item count to site data
site_plots <- left_join(site_plots, item_count, by="ID")

#Total item count of all IDs per survey event
item_total <- site_plots %>%
  group_by(ID) %>%
  mutate(ID_sum=sum(mean_prop), #sum the means in order to divide by count
         ID_prop=ID_sum/IDCount)%>%  #overall mean of each item type
  ungroup()
```


```{r load CSIRO code}
#Load data, assign date to date and notes to character
CSIRO_code <- read.csv(file="Data/CSIRO-code.csv", header=T, sep=",")
CSIRO_code$item_name <- as.character(CSIRO_code$item_name)
```


```{r top item by pier}
#top item by pier
top_pier <- item_total %>%
  filter(Habitat=="P") %>%
  mutate(ID_prop2=paste(ID, ID_prop)) %>%
  distinct(ID_prop2, .keep_all=TRUE) %>%
slice_max(ID_prop, n=25) %>%
  arrange(desc(ID_prop)) %>%
  select(ID, Material, ID_prop)
   
#join CSIRO codes for item names
top_pier <- left_join(top_pier, CSIRO_code, by="ID")
```

```{r top item by soft sediment}
#top item by soft sediment
top_ss <- item_total %>%
  filter(Habitat=="SS") %>%
  mutate(ID_prop2=paste(ID, ID_prop)) %>%
  distinct(ID_prop2, .keep_all=TRUE) %>%
slice_max(ID_prop, n=25) %>%
  arrange(desc(ID_prop)) %>%
  select(ID, Material, ID_prop)
   
#join CSIRO codes for item names
top_ss <- left_join(top_ss, CSIRO_code, by="ID")
```




```{r}
#Consolidating the four main material categories

    
 var_change <- c("Brick or Cement","Construction", "Organic", "Unknown", " ")
 site_total$Material[site_total$Material %in% var_change] = "Other"
    

unique(site_total$Material)

```


