---
title: "gllvm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load-libraries, warning=FALSE, message=FALSE, results='hide'}
library(ggplot2) #graphing
library(ggthemes) #graphing templates
library(hrbrthemes) #graphing templates
library(lubridate) #date manipulation
library(forcats) #working with factors
library(tidyverse) #manipulating data
library(knitr) #rmarkdown functions
library(mvabund) #for multivariate stats
library(gllvm) #for multivariate abundance
library(lattice) #fourth corner heatmap
library(corrplot) #co-occurrence matrix
library(gclus) #co-occurrence matrix
library(broom.mixed) #regression tables

#Tutorial
#https://cran.r-project.org/web/packages/gllvm/vignettes/vignette1.html
```
# Site selection
10 sites at 5 locations were surveyed between 3-5 times each during summer 2019-2020.
```{r create-site-select}
#read in processed survey data
data <- read.csv(file="data/all_data.csv", 
                 header=T, sep=",", 
                 fileEncoding="UTF-8-BOM") #removes special characters

data$Date <- dmy(data$Date) 
data$Notes <- as.character(data$Notes)

#keep only unsw surveys and sites with public piers and >2 survey rounds
site_select <- data %>%
  filter(Team=="UNSW" & 
           Location %in% c("DOUBL","WATSO","PARSL","NEUTR","CLIFT")) %>%
  dplyr::select(-Habitat2, -Fragment, -Whole)

#create a df for survey count per site
sampling_dates <- data %>%
  group_by(Location) %>%
  filter(Team=="UNSW") %>%
  count(Date)

#take away n
sampling_dates$n <- NULL

#create a column for survey count per site
sampling_dates <- sampling_dates %>%
  count(Location) %>%
  arrange(desc(n)) %>%
  rename(NumberSurveys=n)

#join survey count to site select
site_select <- left_join(site_select, sampling_dates, by="Location")

site_select <- site_select %>%
  mutate(survey_event = paste(Date, Site,sep=" "))
```

# Debris Types
CSIRO debris classification system was modified to assign a material type and whether fishing or non-fishing related for each debris item.  Dispersiveness was added as a trait using Tangaroa Blue classification system.
```{r load-item-code}
#load item names
item_code <- read.csv(file="data/CSIRO-code.csv", 
                      header=T, sep=",",
                      fileEncoding="UTF-8-BOM") #removes special characters) 

# #filter out material as some item types may have more than one material type 
# #(ex. fishing items, furniture)
# item_code <- item_code %>%
#   dplyr::select(ID, item_name, Fishing, Dispersiveness) 

site_select <- left_join(site_select, item_code, by="ID", suffix=c("", ".code")) 

#state variables to rename (strict)
var_change <- c("Ceramic", "Undetermined", "E-Waste", "Paper", "Construction", "Organic", "Unknown", "Timber", "Foam", "Brick or Cement", "Rubber", "Cloth", "")

# #state variables to rename (loose)
# var_change <- c("Ceramic", "Undetermined", "Paper", "Construction", "Organic", "Unknown", "Foam", "") 

#change Material to character
site_select$Material <- as.character(site_select$Material) 

#set lesser used categories to 'Other'
site_select$Material[site_select$Material %in% var_change] = "Other" 
```

# Site Attributes
Each location included a 25m transect at a pier and an adjacent 25m soft sediment site in an area of boat traffic.  Distance from harbour mouth was calculated using ArcGIS Cost Distance analysis.
```{r load-site-attributes}
#load site coords and distance from harbour mouth
site_attr <- read.csv(file="data/site_attributes.csv", header=T, sep=",", fileEncoding="UTF-8-BOM")

#add distance to harbour mouth to site_select data
site_attr <- site_attr %>%
  dplyr::select(Site, CostDis) %>% #select only distance to harbour mouth for join
  mutate(DistKm = CostDis/100) #distance to harbour mouth
  
#join distance to mouth to main dataset
site_select <- left_join(site_select, site_attr, by="Site") 
  
```
# Fourth corner model set-up
Creating dataframes for abundance of debris types, environmental variables, and traits of each debris type.
```{r gllvm-set-up, warning=FALSE, message=FALSE, results="hide"}
#Abundance
  #create abundance dataframe with ID as columns and survey events as rows
  abund_long <- site_select %>%
    mutate(survey_event=paste0(Date, Site)) %>% #create column for unique survey event
    group_by(survey_event, ID) %>% 
    summarise(sum=sum(Total)) %>% #total of each item type per survey event
    dplyr::select(survey_event, ID, sum) %>%
  group_by(ID) %>%
  filter(sum>0) %>% #filter IDs that have no values across all surveys
  ungroup()
  
  #convert to data wide
  abund <- spread(abund_long, ID, sum) 
  
  #replace na values with 0
  abund[is.na(abund)] <- 0
  
  #convert survey event to rowname
  length(unique(abund$survey_event)) == nrow(abund) #checking for duplicates
  abund <- column_to_rownames(abund, var="survey_event") #create rownames from survey event

#Environment
  #create survey event and filter zero sum surveys, 
    env <- site_select %>%
    mutate(survey_event=paste0(Date, Site)) %>%
      group_by(survey_event) %>%
      mutate(sum=sum(Total)) %>%
      filter(sum>0) %>%
      dplyr::select(survey_event, Habitat, DistKm) %>%
      distinct()
    
  #convert survey event to rowname
  length(unique(env$survey_event)) == nrow(env) #checking for duplicates
  env <- column_to_rownames(env, var="survey_event") #create rownames from survey event
  #convert Habitat to numeric
  env$Habitat <- as.factor(env$Habitat) #convert character to factor with 2 levels
  env$Habitat <- as.numeric(env$Habitat) #convert factor to numeric
  # #convert Coast to numeric
  # env$Coast <- as.factor(env$Coast) #convert character to factor with 2 levels
  # env$Coast <- as.numeric(env$Coast) #convert factor to numeric

#Traits
  
  #create column of usable IDs to join
  IDs <- abund %>%
    gather(ID, sum) %>%
    dplyr::select(ID) %>%
    unique()
  #create trait df with ID type as row and columns are fishing and dispersiveness
  TR <- site_select %>%
    group_by(ID) %>%
    summarise(sum=sum(Total)) %>%
    left_join(item_code, by="ID") %>%
    dplyr::select(Material, ID, Dispersiveness, Fishing) %>%
    inner_join(IDs, by="ID") %>% #remove items with 0 count
    column_to_rownames(var="ID") #create rownames from ID
  
  #state variables to rename (strict)
  var_change <- c("Ceramic", "Undetermined", "E-Waste", "Paper", "Construction", "Organic", "Unknown", "Timber", "Foam", "Brick or Cement", "Rubber", "Cloth", "")
  
#   #state variables to rename (loose)
# var_change <- c("Ceramic", "Undetermined", "Paper", "Construction", "Organic", "Unknown", "Foam", "Rubber", "Timber", "E-waste", "") 
  
  #change Material to character
  TR$Material <- as.character(TR$Material) 
  
  #set lesser used categories to 'Other'
  TR$Material[TR$Material %in% var_change] = "Other"

  TR$Dispersiveness <- as.factor(TR$Dispersiveness) #convert character to factor with 4 levels
  TR$Fishing <- as.factor(TR$Fishing) #convert character to factor with 2 levels
  TR$Material <- as.factor(TR$Material) #convert character to factor with 4 levels

y <- as.matrix(abund)
X <- scale(as.matrix(env))
TR <- TR
```

```{r wide-format, warning=FALSE, message=FALSE, results="hide"}
# # Model without predictors:
# gllvm(y, family = "negative.binomial")
# # Model where environmental variables, that is, all variables in X are included:
# gllvm(y, X, family = "negative.binomial")
# # Fourth corner model, where all main effects for environmental variables in X and
# # all interactions between variables in X and variables in TR are included:
# fit_wide <- gllvm(y, X, TR, family = "negative.binomial")
# 
# summary(fit_wide)
# 
# plot(fit_wide)
```
```{r long-format, warning=FALSE, message=FALSE, results="hide"}
yX <- reshape(data.frame(cbind(y, X)), direction = "long", varying =
                colnames(y), v.names = "y", timevar = "sp")
#id is the site (1:20), sp is the debris ID, y is the count
TR2 <- data.frame(sp = 1:62, TR)
datalong <- merge(yX, TR2, by = "sp")
```
# Fourth corner model
Coefficient plot shows estimated coefficients (ticks) and their 95% confidence interval (lines). that there are a few significant variables including Habitat affected by fishing, habitat on its own, and the dispersiveness category of unencumbered drift.  

```{r model-long}
#model with environmental variables as predictors
fit_env <- gllvm(y, X, family="negative.binomial", 
                 formula = y ~ (0 + DistKm + Habitat))
coefplot(fit_env)

# Model with environmental and trait variables as predictors
fit_allfixed <- gllvm(data=datalong, 
                      formula = y ~ (0 + Habitat + DistKm) + 
                        (Habitat + DistKm) : (Material + Dispersiveness + Fishing),
                      family = "negative.binomial")
coefplot(fit_allfixed)

anova(fit_env, fit_allfixed)

summary(fit_allfixed)

?levelplot

```
# Latent variables
The latent variables in gllvm would be correlation between types of rubbish (where you either have rubbish type as a categorical predictor or as separate columns in the response matrix). You would expect e.g. correlation between types of rubbish that may be generated from similar activities- e.g. beach parties, fishing etc. e.g. bottles and bottle caps and plastic wrappers may have a relationship, all coming from beach parties. Correlation can also come from missing predictors in the model. -- Eve
```{r latent-variables}
# #biplot - clusters of debris items with numbers representing sites
# ordiplot(fit_allfixed, biplot=TRUE)
# 
# fit_env <- gllvm(y, X, family="negative.binomial") #abundance + env traits
# cr <- getResidualCor(fit_env) #residual correlation matrix
# 
# #correlation matrix for debris types
# corrplot(cr[order.single(cr), 
#             order.single(cr)], 
#          diag=FALSE, #whether display coefs on the principal diagonal
#          type="lower", #displays lower triangle matrix
#          method="circle", #visualization method
#          tl.cex = 0.5, tl.srt=45, tl.col="red")

```

# Fourth corner heatmap
Here we see that nothing is significantly interacting with distance from harbour mouth or coast, however fishing interacts with habitat significantly. Brighter colours represent stronger associations, red are positive associations and blue are negative.
```{r 4th-corner-heatmap}
fourth <- fit_allfixed$fourth.corner
a <- max( abs(fourth) )
colort <- colorRampPalette(c("blue", "white", "red"))
plot.4th <- levelplot((as.matrix(fourth)), 
                      xlab = "Environmental Variables", 
                      ylab = "Debris traits",
                      col.regions = colort(100), #use defined colour palette
                      cex.lab = 1.3, #specify size of the label text
                      at = seq(-a, a, length = 100), #makes heat scale equal both sides
                      scales = list(x = list(rot = 45))) #rotate text for clarity


plot.4th
```
