---
title: "results_figures"
author: "Brie Sherow"
date: "04/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries, warning=FALSE, message=FALSE, results='hide'}
library(ggplot2) #graphing
library(ggthemes) #graphing templates
library(hrbrthemes) #graphing templates
library(lubridate) #date manipulation
library(forcats) #working with factors
library(tidyverse) #manipulating data
library(knitr) #rmarkdown functions
#library(kableExtra) #table layouts
library(magick) #image processing
library(stats) #R stats functions
library(broom) #create summaries from stats objects
library(car) #lm regression
library(MASS) #stats
library(lme4) #glmer function
library(DHARMa) #testing model diagnostics
library(glmmTMB) #fit zero-inflated negative binomial
# library(mvabund) #for multivariate stats
# library(gllvm) #for multivariate abundance
library(lattice) #fourth corner heatmap
library(vegan)
```

# Site selection
10 sites at 5 locations were surveyed between 3-5 times each during summer 2019-2020.
```{r create-site-select}
#read in processed survey data
data <- read.csv(file="data/all_data.csv", 
                 header=T, sep=",", 
                 fileEncoding="UTF-8-BOM") #removes special characters

data$Date <- dmy(data$Date) 
data$Notes <- as.character(data$Notes)

#keep only unsw surveys and sites with public piers and >2 survey rounds
site_select <- data %>%
  filter(Team=="UNSW" & 
           Location %in% c("DOUBL","WATSO","PARSL","NEUTR","CLIFT")) %>%
  dplyr::select(-Habitat2, -Fragment, -Whole)

#create a df for survey count per site
sampling_dates <- data %>%
  group_by(Location) %>%
  filter(Team=="UNSW") %>%
  count(Date)

#take away n
sampling_dates$n <- NULL

#create a column for survey count per site
sampling_dates <- sampling_dates %>%
  count(Location) %>%
  arrange(desc(n)) %>%
  rename(NumberSurveys=n)

#join survey count to site select
site_select <- left_join(site_select, sampling_dates, by="Location")

site_select <- site_select %>%
  mutate(survey_event = paste(Date, Site,sep=" "))
```

# Debris types
CSIRO debris classification system was modified to assign a material type and whether they are fishing or non-fishing related for each debris item.  Dispersiveness was added as a trait using Tangaroa Blue classification system.
```{r load-item-code}
#load item names
item_code <- read.csv(file="data/CSIRO-code.csv", 
                      header=T, sep=",",
                       fileEncoding="UTF-8-BOM") #removes special characters

#filter out material as some item types may have more than one material type 
#(ex. fishing items, furniture)
item_code <- item_code %>%
  dplyr::select(ID, item_name, Material, Dispersiveness, Fishing, Label) 

site_select <- left_join(site_select, item_code, by="ID", suffix=c("", ".code"))

site_select <- site_select %>%
  mutate(item = paste(Material, item_name, sep=": "))
```

# Site attributes
Each location included a 25m transect at a pier and an adjacent 25m soft sediment site in an area of boat traffic.  Distance from harbour mouth was calculated using ArcGIS Cost Distance analysis.
```{r load-site-attributes}
#load site coords and distance from harbour mouth
site_attr <- read.csv(file="data/site_attributes.csv", header=T, sep=",", fileEncoding="UTF-8-BOM")

#add distance to harbour mouth to site_select data
site_attr <- site_attr %>%
  dplyr::select(Site, CostDis, Coast) %>% #select only distance to harbour mouth for join
  mutate(DistKm = CostDis/100) #distance to harbour mouth
  
#join distance to mouth to main dataset
site_select <- left_join(site_select, site_attr, by="Site") 
```

```{r high-level}
t <- site_select %>%
  group_by(Habitat) %>%
  summarise(sum=sum(Total))
```


```{r results-map, fig.cap="Fig 1: Mean debris count and proportion per habitat", out.width = '100%'}
knitr::include_graphics("images/Sites.svg")
```


```{r top-items-lollipop}
top_lol <- site_select 

#state variables to rename
top20 <- c("F2", "F6", "G1", "M10", "F7", "Z2", "Z5", "C1", "BP3", "S2", "H10", "G4", "R5", "T5", "F3", "S6", "C2", "Z4", "S7", "H6") 

get_mean <- site_select %>%
  filter(ID %in% top20) %>%
  group_by(Habitat, Label) %>%
  summarise(sum=sum(Total)) %>%
  spread(Habitat, sum) %>%
  rowwise() %>% 
  mutate( mean = mean(c(P,SS) ))

top_lol <- site_select %>%
  filter(ID %in% top20) %>%
  group_by(Habitat, ID, Label) %>%
  summarise(sum=sum(Total)) %>%
  left_join( get_mean, by="Label") %>%
  mutate(Label=as.factor(Label),
  Label=fct_reorder( Label, mean, .desc=TRUE)) %>%
  ungroup()

HabitatLabs <- c(P = "Pier", SS = "Sediment")
top_lol$Habitat[top_lol$Habitat == "P"] <-"Pier"
top_lol$Habitat[top_lol$Habitat == "SS"] <-"Sediment"

p_top_lol <- top_lol %>% 
  ungroup() %>%
  mutate(Label=as.factor(Label),
  Label=fct_reorder(Label, mean)) %>%
  dplyr::select(-P, -SS) %>%
  ggplot( aes(x=sum, y=Label)) +
  scale_x_log10() +
  geom_linerange( aes(xmin=0, xmax=sum, y=Label, colour=Habitat), position=position_dodge(.5)) +
  geom_point( aes(colour=Habitat), position=position_dodge(.5), size=2) +
   # labs(title="Most abundant debris items",
   #     subtitle="Total counts by habitat (log10)",
   #     x="Total count", y="Debris item types", tag="C") +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        panel.background = element_blank()) 


p_top_lol

# ggsave(filename = "images/p_toplol.svg", plot = p_top_lol, device="svg")

svg("images/p_toplol.svg")
plot(p_top_lol)
dev.off()
```

## GLMM negative binomial model
I aggregated debris totals by survey date and location and used a negative binomial GLMM model to fit the data.  Debris totals are predicted by habitat, distance from harbour mouth, and aspect (north or south coast).  Location and Date are random effects.  

```{r glmm-hab, warning=FALSE, message=FALSE}

#create df to use for models
mod_hab <- site_select %>%
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) %>%
  ungroup() %>%
  mutate(Location = as.factor(Location),
         Habitat = as.factor(Habitat))


m_hab  <- glmmTMB(Total ~ #Debris count
                 Habitat * DistKm + #predicted by Habitat
                 (1|Location) + (1|Date), #date random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_hab)

# Predictors are: Habitat + DistKm

nd_hab <- expand.grid(Habitat = unique(mod_hab$Habitat),
                  DistKm = seq(from = min(mod_hab$DistKm),
                               to = max(mod_hab$DistKm),
                               length=1000),
                  Location = NA,
                  Date = NA)

pred_hab <- predict(object = m_hab,
                         newdata = nd_hab,
                         se.fit = T,
                         re.form = NA,
                         type="response")
                    

nd_hab$Total <- pred_hab$fit
nd_hab$SE_upper <- pred_hab$fit + pred_hab$se.fit
nd_hab$SE_lower <- pred_hab$fit - pred_hab$se.fit

#plot the predicted means
p_hab <- ggplot(nd_hab, aes(y=Total, x=DistKm)) + 
  geom_line(aes(col=Habitat)) +
  labs(title="Debris abundance predictions",
       subtitle="By habitat and distance from harbour mouth",
       x="Distance from harbour mouth (km)", y="Total debris items", tag="") +
  geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) 
    # theme(legend.position = "none",
    #     panel.grid = element_blank(),
    #     axis.title = element_blank(),
    #     axis.text = element_blank(),
    #     panel.background = element_blank()) 

p_hab

# ggsave(filename = "images/p_hab.svg", plot = p_hab, device="svg", height=7, width=12, units="in")

svg("images/p_hab.svg")
plot(p_hab, height=7, width=12, units="in")
dev.off()

```

```{r glm-fish, warning=FALSE, message=FALSE}
#create df to use for models
mod_fish <- site_select %>%
  filter(Fishing=="Fishing") %>% #fishing line
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) %>%
  ungroup() %>%
              mutate(Location = as.factor(Location),
         Habitat = as.factor(Habitat))

#GLM negative binomial with two variables and zero inflation
m_fish  <- glmmTMB(Total ~ #Debris count
                 Habitat * DistKm + #predicted by distance from harbour mouth
                   # Coast + #predicted by north or south coast
                 (1|Location) + (1|Date), #site random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_fish) 

# Predictors for fishing related debris are: Habitat + DistKm

nd_fish <- expand.grid(Habitat = unique(mod_fish$Habitat),
                  DistKm = seq(from = min(mod_fish$DistKm),
                               to = max(mod_fish$DistKm),
                               length=1000),
                  Location = NA,
                  Date = NA)

pred_fish <- predict(object = m_fish,
                         newdata = nd_fish,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_fish$Total <- pred_fish$fit
nd_fish$SE_upper <- pred_fish$fit + pred_fish$se.fit
nd_fish$SE_lower <- pred_fish$fit - pred_fish$se.fit

p_fish <- ggplot(nd_fish, aes(y=Total, x=DistKm)) + 
  geom_line(aes(col=Habitat)) +
  geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        panel.background = element_blank()) 

p_fish

# ggsave(filename = "images/p_fish.svg", plot = p_fish, device="svg", height=7, width=7, units="in")

svg("images/p_fish.svg")
plot(p_fish, height=7, width=7, units="in")
dev.off()

   # labs(title="Debris abundance predictions: Fishing related",
    #    subtitle="By habitat and distance from harbour mouth",
       # x="Distance from harbour mouth (km)", y="Total debris items", tag="") 

```
## Testing non-fishing related debris
```{r glm-nonfish, warning=FALSE, message=FALSE}
#create df to use for models
mod_nonfish <- site_select %>%
  filter(Fishing=="Non-Fishing") %>% #nonfishing debris
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) %>%
  ungroup() %>%
              mutate(Location = as.factor(Location),
         Habitat = as.factor(Habitat))

m_nonfish  <- glmmTMB(Total ~ #Debris count
                 Habitat + #predicted by Habitat
                 (1|Location) + (1|Date), #site random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_nonfish)

# Predictors for non-fishing related debris are: Habitat

nd_nonfish <- expand.grid(Habitat = unique(mod_nonfish$Habitat),
                          # DistKm = seq(from = min(mod_nonfish$DistKm),
                          #      to = max(mod_nonfish$DistKm),
                          #      length=1000),
                  Location = NA,
                  Date = NA)


pred_nonfish <- predict(object = m_nonfish,
                         newdata = nd_nonfish,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_nonfish$Total <- pred_nonfish$fit
nd_nonfish$SE_upper <- pred_nonfish$fit + pred_nonfish$se.fit
nd_nonfish$SE_lower <- pred_nonfish$fit - pred_nonfish$se.fit


p_nonfish <- ggplot(nd_nonfish, aes(y=Total, x=Habitat)) + 
  geom_col(aes(fill=Habitat)) +
  geom_errorbar(aes(ymin=SE_lower, ymax=SE_upper, width=0.2), position=position_dodge(width=0.90)) +
   theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        panel.background = element_blank()) 

p_nonfish

# ggsave(filename = "images/p_nonfish.svg", plot = p_nonfish, device="svg", height=7, width=7, units="in")

svg("images/p_nonfish.svg")
plot(p_nonfish, height=7, width=7, units="in")
dev.off()

```
```{r glm-F2, warning=FALSE, message=FALSE}
#create df of fishing line to use for models
mod_F2 <- site_select %>%
  filter(ID=="F2") %>% #fishing line
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) %>%
  ungroup() %>%
              mutate(Location = as.factor(Location),
         Habitat = as.factor(Habitat))

#GLM negative binomial with two variables and zero inflation
m_F2  <- glmmTMB(Total ~ #Debris count
                 Habitat * DistKm + #distance from harbour mouth
                 (1|Location) + (1|Date), #site random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_F2) 
# Predictors are: Habitat, DistKm

nd_F2 <- expand.grid(Habitat = unique(mod_F2$Habitat),
                          DistKm = seq(from = min(mod_F2$DistKm),
                               to = max(mod_F2$DistKm),
                               length=1000),
                  Location = NA,
                  Date = NA)

pred_F2 <- predict(object = m_F2,
                         newdata = nd_F2,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_F2$Total <- pred_F2$fit
nd_F2$SE_upper <- pred_F2$fit + pred_F2$se.fit
nd_F2$SE_lower <- pred_F2$fit - pred_F2$se.fit

#plot the predicted means

p_F2 <- ggplot(nd_F2, aes(y=Total, x=DistKm)) +
  geom_line(aes(col=Habitat)) +
  geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) 
   # theme(legend.position = "none",
   #      panel.grid = element_blank(),
   #      axis.title = element_blank(),
   #      axis.text = element_blank(),
   #      panel.background = element_blank()) 

p_F2

# ggsave(filename = "images/p_F2.svg", plot = p_F2, device="svg", height=7, width=7, units="in")

# svg("images/p_F2.svg")
# plot(p_F2, height=7, width=7, units="in")
# dev.off()

```
## Testing glass bottles debris type
```{r glm-btl, warning=FALSE, message=FALSE}
#create df of glass bottles to use for models
mod_btl <- site_select %>%
  filter(ID=="G1") %>% #glass bottles
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) %>%
  ungroup() %>%
              mutate(Location = as.factor(Location),
         Habitat = as.factor(Habitat))

#GLM negative binomial with two variables and zero inflation
# m_btl1  <- glmmTMB(Total ~ #Debris count
#                  Habitat*DistKm + #predicted by Habitat
#                  (1|Location) + (1|Date), #site/date random effect
#                family=nbinom1(), #negative binomial to deal with count data and zeros
#                data = mod_btl)

m_btl  <- glmmTMB(Total ~ #Debris count
                 Habitat + #predicted by Habitat
                 (1|Location) + (1|Date), #site/date random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_btl)

# Predictors for glass bottles are: Habitat, Coast

nd_btl <- expand.grid(Habitat = unique(mod_btl$Habitat),
                  Location = NA,
                  Date = NA)

pred_btl <- predict(object = m_btl,
                         newdata = nd_btl,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_btl$Total <- pred_btl$fit
nd_btl$SE_upper <- pred_btl$fit + pred_btl$se.fit
nd_btl$SE_lower <- pred_btl$fit - pred_btl$se.fit

#plot the predicted means

p_btl <- ggplot(nd_btl, aes(y=Total, x=Habitat)) +
  geom_col(aes(fill=Habitat)) +
  geom_errorbar(aes(ymin=SE_lower, ymax=SE_upper, width=0.2), position=position_dodge(width=0.90)) 
   # theme(legend.position = "none",
   #      panel.grid = element_blank(),
   #      axis.title = element_blank(),
   #      axis.text = element_blank(),
   #      panel.background = element_blank()) 

p_btl

# ggsave(filename = "images/p_btl.svg", plot = p_btl, device="svg", height=7, width=7, units="in")

# svg("images/p_btl.svg")
# plot(p_btl, height=7, width=7, units="in")
# dev.off()

```

# glm plastic
```{r glm-plastic, warning=FALSE, message=FALSE}
#create df of all plastic materials
mod_plas <- site_select %>%
  filter(Material=="Plastic") %>% 
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) 

#GLM negative binomial with two variables and zero inflation
m_plas  <- glmmTMB(Total ~ #Debris count
                 Habitat*DistKm + #predicted by Habitat
                 (1|Location) + (1|Date), #site/date random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_plas) 

# Predictors for plastic are: Habitat, DistKm

nd_plas <- expand.grid(Habitat = unique(mod_plas$Habitat),
                          DistKm = seq(from = min(mod_plas$DistKm),
                               to = max(mod_plas$DistKm),
                               length=1000),
                  Location = NA,
                  Date = NA)

pred_plas <- predict(object = m_plas,
                         newdata = nd_plas,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_plas$Total <- pred_plas$fit
nd_plas$SE_upper <- pred_plas$fit + pred_plas$se.fit
nd_plas$SE_lower <- pred_plas$fit - pred_plas$se.fit

#plot the predicted means

p_plas <- ggplot(nd_plas, aes(y=Total, x=DistKm)) +
  geom_line(aes(col=Habitat)) +
  geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) 
   # theme(legend.position = "none",
   #      panel.grid = element_blank(),
   #      axis.title = element_blank(),
   #      axis.text = element_blank(),
   #      panel.background = element_blank()) 

p_plas

# ggsave(filename = "images/p_plas.svg", plot = p_plas, device="svg", height=7, width=7, units="in")

# svg("images/p_plas.svg")
# plot(p_plas, height=7, width=7, units="in")
# dev.off()

```
# glm glass
```{r glm-glsmat, warning=FALSE, message=FALSE}
#create df of all misc materials
mod_gls <- site_select %>%
  filter(Material=="Glass") %>% #glass
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) 


m_gls  <- glmmTMB(Total ~ #Debris count
                 Habitat + #predicted by Habitat
                 (1|Location) + (1|Date), #site/date random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_gls)

# Predictors for 'other' materials are: Habitat, Coast

nd_gls <- expand.grid(Habitat = unique(mod_gls$Habitat),
                  Location = NA,
                  Date = NA)

pred_gls <- predict(object = m_gls,
                         newdata = nd_gls,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_gls$Total <- pred_gls$fit
nd_gls$SE_upper <- pred_gls$fit + pred_gls$se.fit
nd_gls$SE_lower <- pred_gls$fit - pred_gls$se.fit

#plot the predicted means

p_gls <- ggplot(nd_gls, aes(y=Total, x=Habitat)) +
  geom_col(aes(fill=Habitat)) +
  geom_errorbar(aes(ymin=SE_lower, ymax=SE_upper, width=0.2), position=position_dodge(width=0.90)) +
  ylim(0, 30) +
   theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        panel.background = element_blank())

p_gls

# ggsave(filename = "images/p_gls.svg", plot = p_gls, device="svg", height=7, width=7, units="in")

svg("images/p_gls.svg")
plot(p_gls, height=7, width=7, units="in")
dev.off()

```
# glm metal
```{r glm-metmat, warning=FALSE, message=FALSE}
#create df of all misc materials
mod_met <- site_select %>%
  filter(Material=="Metal") %>% #glass
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) 

# #GLM negative binomial with two variables and zero inflation
m_met  <- glmmTMB(Total ~ #Debris count
                 Habitat + #predicted by Habitat
                 (1|Location) + (1|Date), #site/date random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_met)

# Predictors for 'other' materials are: Habitat, Coast

nd_met <- expand.grid(Habitat = unique(mod_met$Habitat),
                  Location = NA,
                  Date = NA)

pred_met <- predict(object = m_met,
                         newdata = nd_met,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_met$Total <- pred_met$fit
nd_met$SE_upper <- pred_met$fit + pred_met$se.fit
nd_met$SE_lower <- pred_met$fit - pred_met$se.fit

#plot the predicted means

p_met <- ggplot(nd_met, aes(y=Total, x=Habitat)) +
  geom_col(aes(fill=Habitat)) +
  geom_errorbar(aes(ymin=SE_lower, ymax=SE_upper, width=0.2), position=position_dodge(width=0.90)) +
  ylim(0, 30) 
   # theme(legend.position = "none",
   #      panel.grid = element_blank(),
   #      axis.title = element_blank(),
   #      axis.text = element_blank(),
   #      panel.background = element_blank())

p_met

# svg("images/p_met.svg")
# plot(p_met, height=7, width=7, units="in")
# dev.off()

# ggsave(filename = "images/p_met.svg", plot = p_met, device="svg", height=7, width=7, units="in")
```

## Ordination
```{r vegan}

#Abundance data
  #Filter for top items
top10 <- site_select %>%
  group_by(Label) %>%
  summarise(sum=sum(Total)) %>%
  arrange(desc(sum)) %>%
  top_n(10, sum)

#create vector from top 10 items
top10 <- top10$Label
top10 <- unlist(top10)
site_select_top10 <- site_select %>%
  filter(Label %in% top10)

  #create abundance dataframe with ID as columns and survey events as rows
  abund_long <- site_select_top10 %>%
    mutate(survey_event=paste0(Date, Site)) %>% #create column for unique survey event
    group_by(survey_event, Location, Habitat, Date, Label) %>% 
    summarise(sum=sum(Total)) %>% #total of each item type per survey event
  group_by(Label) %>%
  filter(sum>0) %>% #filter IDs that have no values across all surveys
  ungroup()
  
  #convert to data wide
  abund <- spread(abund_long, Label, sum) 
  
  #replace na values with 0
  abund[is.na(abund)] <- 0
  
  #convert survey event to rowname
  length(unique(abund$survey_event)) == nrow(abund) #checking for duplicates
  abund <- column_to_rownames(abund, var="survey_event") #create rownames from survey event
  
  com <- abund[4:13]
  
  ord <- metaMDS(com)
  
  ord$stress
  # 0.1419583
  
  data.scores <- as.data.frame(scores(ord)) #create df of ord points
  data.scores$Location <- abund$Location
  data.scores$Habitat <- abund$Habitat
  data.scores <- data.scores %>%
    mutate(Habitat = as.factor(Habitat),
           Location = as.factor(Location))
  
  species.scores <- as.data.frame(scores(ord, "species"))
  species.scores$item <- rownames(species.scores) #create item id column
  
#create ellipsis
#add group column to data.scores
NMDS <- data.frame(NMDS1 = ord$points[,1], NMDS2 = ord$points[,2],group=data.scores$Habitat)
 veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
  {
    theta <- (0:npoints) * 2 * pi/npoints
    Circle <- cbind(cos(theta), sin(theta))
    t(center + scale * t(Circle %*% chol(cov)))
  }
  df_ell <- data.frame()
  for(g in levels(NMDS$group)){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$group==g,],
                    veganCovEllipse(cov.wt(cbind(NMDS1,NMDS2),wt=rep(1/length(NMDS1),length(NMDS1)))$cov,center=c(mean(NMDS1),mean(NMDS2)))))
                    ,group=g))
  }

  p_ord <- ggplot() +
  # geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2, label=item)) +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), 
            size=1, linetype=2) +
  geom_point(data=data.scores, 
             aes(x=NMDS1, y=NMDS2, colour=Habitat), alpha=0.8) + 
  coord_equal() +
   theme(legend.position = "none",
        panel.background = element_blank())
  
  p_ord

# svg("images/p_ord.svg", height=7, width=11)
# plot(p_ord, height=7, width=11, units="in")
# dev.off()

  p_ord_txt <- ggplot() +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), 
            size=1, linetype=2) +
  geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2, label=item)) +
  geom_point(data=data.scores, 
             aes(x=NMDS1, y=NMDS2, colour=Habitat), alpha=0.8) + 
  coord_equal() +
   theme(legend.position = "none",
        panel.background = element_blank())
  
p_ord_txt

# svg("images/p_ord_txt.svg", height=7, width=11)
# plot(p_ord_txt, height=7, width=11, units="in")
# dev.off()

```

```{r vegan-20}

#Abundance data
#   #Filter for top items
# top20 <- site_select %>%
#   group_by(Label) %>%
#   summarise(sum=sum(Total)) %>%
#   arrange(desc(sum)) %>%
#   top_n(20, sum)
# 
# #create vector from top 20 items
# top20 <- top20$Label
# top20 <- unlist(top20)
site_select_top20 <- site_select %>%
  filter(ID %in% top20)

  #create abundance dataframe with ID as columns and survey events as rows
  abund_long <- site_select_top20 %>%
    mutate(survey_event=paste0(Date, Site)) %>% #create column for unique survey event
    group_by(survey_event, Location, Habitat, Date, Label) %>% 
    summarise(sum=sum(Total)) %>% #total of each item type per survey event
  group_by(Label) %>%
  filter(sum>0) %>% #filter IDs that have no values across all surveys
  ungroup()
  
  #convert to data wide
  abund <- spread(abund_long, Label, sum) 
  
  #replace na values with 0
  abund[is.na(abund)] <- 0
  
  #convert survey event to rowname
  length(unique(abund$survey_event)) == nrow(abund) #checking for duplicates
  abund <- column_to_rownames(abund, var="survey_event") #create rownames from survey event
  
  com <- abund[4:23]
  
  ord <- metaMDS(com)
  
  data.scores <- as.data.frame(scores(ord)) #create df of ord points
  data.scores$Location <- abund$Location
  data.scores$Habitat <- abund$Habitat
  data.scores <- data.scores %>%
    mutate(Habitat = as.factor(Habitat),
           Location = as.factor(Location))
  
  species.scores <- as.data.frame(scores(ord, "species"))
  species.scores$item <- rownames(species.scores) #create item id column

#create ellipsis
#add group column to data.scores
NMDS <- data.frame(NMDS1 = ord$points[,1], NMDS2 = ord$points[,2],group=data.scores$Habitat)
 veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
  {
    theta <- (0:npoints) * 2 * pi/npoints
    Circle <- cbind(cos(theta), sin(theta))
    t(center + scale * t(Circle %*% chol(cov)))
  }
  df_ell <- data.frame()
  for(g in levels(NMDS$group)){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$group==g,],
                    veganCovEllipse(cov.wt(cbind(NMDS1,NMDS2),wt=rep(1/length(NMDS1),length(NMDS1)))$cov,center=c(mean(NMDS1),mean(NMDS2)))))
                    ,group=g))
  }

  p_ord <- ggplot() +
  # geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2, label=item)) +
  geom_point(data=data.scores, 
             aes(x=NMDS1, y=NMDS2, colour=Habitat), alpha=0.8) + 
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), 
            size=1, linetype=2) +
  coord_equal() +
   theme(legend.position = "none",
        panel.background = element_blank())
  
  p_ord

svg("images/p_ord.svg", height=7, width=11)
plot(p_ord, height=7, width=11, units="in")
dev.off()

  p_ord_txt <- ggplot() +
  geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2, label=item)) +
  geom_point(data=data.scores, 
             aes(x=NMDS1, y=NMDS2, colour=Habitat), alpha=0.8) + 
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), 
            size=1, linetype=2) +
  coord_equal() +
   theme(legend.position = "none",
        panel.background = element_blank())
  
  p_ord_txt

# svg("images/p_ord_txt.svg", height=7, width=11)
# plot(p_ord_txt, height=7, width=11, units="in")
# dev.off()

```
## Fourth corner analysis
Determining the significance of environmental traits and debris attributes.
```{r gllvm-set-up, warning=FALSE, message=FALSE, results="hide"}
#Abundance
  #create abundance dataframe with ID as columns and survey events as rows
  abund_long <- site_select %>%
    mutate(survey_event=paste0(Date, Site)) %>% #create column for unique survey event
    group_by(survey_event, ID) %>% 
    summarise(sum=sum(Total)) %>% #total of each item type per survey event
    dplyr::select(survey_event, ID, sum) %>%
  group_by(ID) %>%
  filter(sum>0) %>% #filter IDs that have no values across all surveys
  ungroup()
  
  #convert to data wide
  abund <- spread(abund_long, ID, sum) 
  
  #replace na values with 0
  abund[is.na(abund)] <- 0
  
  #convert survey event to rowname
  length(unique(abund$survey_event)) == nrow(abund) #checking for duplicates
  abund <- column_to_rownames(abund, var="survey_event") #create rownames from survey event

#Environment
  #create survey event and filter zero sum surveys, 
    env <- site_select %>%
    mutate(survey_event=paste0(Date, Site)) %>%
      group_by(survey_event) %>%
      mutate(sum=sum(Total)) %>%
      filter(sum>0) %>%
      dplyr::select(survey_event, Habitat, DistKm, Location) %>%
      distinct()
    
  #convert survey event to rowname
  length(unique(env$survey_event)) == nrow(env) #checking for duplicates
  env <- column_to_rownames(env, var="survey_event") #create rownames from survey event
  #convert Habitat to numeric
  env$Habitat <- as.factor(env$Habitat) #convert character to factor with 2 levels
  env$Habitat <- as.numeric(env$Habitat) #convert factor to numeric
  #convert Location to numeric
  env$Location <- as.factor(env$Location) #convert character to factor with 2 levels
  env$Location <- as.numeric(env$Location) #convert factor to numeric

#Traits
  
  #create column of usable IDs to join
  IDs <- abund %>%
    gather(ID, sum) %>%
    dplyr::select(ID) %>%
    unique()
  #create trait df with ID type as row and columns are fishing and dispersiveness
  TR <- site_select %>%
    group_by(ID) %>%
    summarise(sum=sum(Total)) %>%
    left_join(item_code, by="ID") %>%
    dplyr::select(Material, ID, Dispersiveness, Fishing) %>%
    inner_join(IDs, by="ID") %>% #remove items with 0 count
    column_to_rownames(var="ID") #create rownames from ID
  
  #state variables to rename
  var_change <- c("Ceramic", "Undetermined", "E-Waste", "Paper", "Construction", "Organic", "Unknown", "Timber", "Foam", "Brick or Cement", "Rubber", "Cloth", "") 
  
  #change Material to character
  TR$Material <- as.character(TR$Material) 
  
  #set lesser used categories to 'Other'
  TR$Material[TR$Material %in% var_change] = "Other" 

  TR$Dispersiveness <- as.factor(TR$Dispersiveness) #convert character to factor with 4 levels
  TR$Fishing <- as.factor(TR$Fishing) #convert character to factor with 2 levels
  TR$Material <- as.factor(TR$Material) #convert character to factor with 4 levels

y <- as.matrix(abund)
X <- scale(as.matrix(env))
TR <- TR

#Data long
yX <- reshape(data.frame(cbind(y, X)), direction = "long", varying =
                colnames(y), v.names = "y", timevar = "sp")
#id is the site (1:20), sp is the debris ID, y is the count
TR2 <- data.frame(sp = 1:62, TR)
datalong <- merge(yX, TR2, by = "sp")
```

Here we see that nothing is interacting significantly with distance to harbour mouth, however both fishing and encumbered drift are interacting with habitat significantly.

# Fourth corner model
Coefficient plot shows estimated coefficients (ticks) and their 95% confidence interval (lines). that there are a few significant variables including Habitat affected by fishing, habitat on its own, and the dispersiveness category of unencumbered drift.  
```{r 4th-corner-heatmap, fig.cap="Figure 5: Fourth corner heatmap"}
# Model with environmental and trait variables as predictors
fit_allfixed <- gllvm(data=datalong, 
                      formula = y ~ (0 + Habitat + DistKm) + 
                        (Habitat + DistKm) * (Material + Dispersiveness + Fishing),
                      family = "negative.binomial")

#call svg command to start the plot
svg(file = "images/coefplot.svg",   # The directory you want to save the file in
    width = 11, # The width of the plot in inches
    height = 7) # The height of the plot in inches
coefplot(fit_allfixed)
dev.off()

t <- fit_allfixed$data

png(file = "images/coefplot.png")  # The directory you want to save the file in
coefplot(fit_allfixed, cex.ylab = .8)
dev.off()

fourth <- fit_allfixed$fourth.corner
a <- max( abs(fourth) )
colort <- colorRampPalette(c("blue", "white", "red"))
plot.4th <- levelplot((as.matrix(fourth)), xlab = "Environmental Variables", 
                      ylab = "Debris traits", col.regions = colort(100), cex.lab = 1.3, 
                      at = seq(-a, a, length = 100), scales = list(x = list(rot = 45)))

#call svg command to start the plot
svg(file = "images/plot4th.svg",   # The directory you want to save the file in
    width = 11, # The width of the plot in inches
    height = 7) # The height of the plot in inches
plot(plot.4th)
dev.off()
```



