---
title: "results_figures"
author: "Brie Sherow"
date: "04/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-libraries, warning=FALSE, message=FALSE, results='hide'}
library(ggplot2) #graphing
library(ggthemes) #graphing templates
library(hrbrthemes) #graphing templates
library(lubridate) #date manipulation
library(forcats) #working with factors
library(tidyverse) #manipulating data
library(knitr) #rmarkdown functions
#library(kableExtra) #table layouts
# library(magick) #image processing
library(stats) #R stats functions
library(broom) #create summaries from stats objects
library(car) #lm regress
library(MASS) #stats
library(lme4) #glmer function
library(DHARMa) #testing model diagnostics
library(glmmTMB) #fit zero-inflated negative binomial
library(mvabund) #for multivariate stats
# library(gllvm) #for multivariate abundance
library(lattice) #fourth corner heatmap
library(vegan)
```

# Site selection
10 sites at 5 locations were surveyed between 3-5 times each during summer 2019-2020.
```{r create-site-select}
#Site selection

  #read in processed survey data
  data <- read.csv(file="data/all_data.csv", 
                   header=T, sep=",", 
                   fileEncoding="UTF-8-BOM") #removes special characters
  
  data$Date <- dmy(data$Date) 
  data$Notes <- as.character(data$Notes)
  
  #keep only unsw surveys and sites with public piers and >2 survey rounds
  site_select <- data %>%
    filter(Team=="UNSW" & 
             Location %in% c("DOUBL","WATSO","PARSL","NEUTR","CLIFT")) %>%
    dplyr::select(-Habitat2, -Fragment, -Whole)
  
  #create a df for survey count per site
  sampling_dates <- data %>%
    group_by(Location) %>%
    filter(Team=="UNSW") %>%
    count(Date)
  
  #take away n
  sampling_dates$n <- NULL
  
  #create a column for survey count per site
  sampling_dates <- sampling_dates %>%
    count(Location) %>%
    arrange(desc(n)) %>%
    rename(NumberSurveys=n)
  
  #join survey count to site select
  site_select <- left_join(site_select, sampling_dates, by="Location")
  
  site_select <- site_select %>%
    mutate(survey_event = paste(Date, Site,sep=" "))

#Debris types
  
  #load item names
  item_code <- read.csv(file="data/CSIRO-code.csv", 
                        header=T, sep=",",
                         fileEncoding="UTF-8-BOM") #removes special characters
  
  #filter out material as some item types may have more than one material type 
  #(ex. fishing items, furniture)
  item_code <- item_code %>%
    dplyr::select(ID, item_name, Material, Dispersiveness, Fishing, Label) 
  
  site_select <- left_join(site_select, item_code, by="ID", suffix=c("", ".code"))
  
  site_select <- site_select %>%
    mutate(item = paste(Material, item_name, sep=": "))

#Site attributes
  
  #load site coords and distance from harbour mouth
  site_attr <- read.csv(file="data/site_attributes.csv", header=T, sep=",", fileEncoding="UTF-8-BOM")
  
  #add distance to harbour mouth to site_select data
  site_attr <- site_attr %>%
    dplyr::select(Site, CostDis, Coast) %>% #select only distance to harbour mouth for join
    mutate(DistKm = CostDis/100) #distance to harbour mouth
  
    #join distance to mouth to main dataset
  site_select <- left_join(site_select, site_attr, by="Site") 
    
  #hab labels
  
  site_select$Habitat <- factor(site_select$Habitat, levels = c("P", "SS"),
                  labels = c("Pier", "Sediment"))
```

```{r high-level}
t <- site_select %>%
  group_by(Habitat) %>%
  summarise(sum=sum(Total))
```


```{r results-map, fig.cap="Fig 1: Mean debris count and proportion per habitat", out.width = '100%'}
knitr::include_graphics("images/Sites.svg")
```


```{r top-items-lollipop}
top_lol <- site_select 

#Abundance data
  #Filter for top items
top20 <- site_select %>%
  group_by(Label) %>%
  summarise(sum=sum(Total)) %>%
  arrange(desc(sum)) %>%
  top_n(20, sum)

#create vector from top 10 items
top20 <- top20$Label
top20 <- unlist(top20)
site_select_top20 <- site_select %>%
  filter(Label %in% top20)

get_mean <- site_select %>%
  filter(Label %in% top20) %>%
  group_by(Habitat, Label) %>%
  summarise(sum=sum(Total)) %>%
  spread(Habitat, sum) %>%
  rowwise() %>% 
  mutate( mean = mean(c(Pier, Sediment) ))

top_lol <- site_select %>%
  filter(Label %in% top20) %>%
  group_by(Habitat, ID, Label) %>%
  summarise(sum=sum(Total)) %>%
  left_join( get_mean, by="Label") %>%
  mutate(Label=as.factor(Label),
  Label=fct_reorder( Label, mean, .desc=TRUE)) %>%
  ungroup()

# HabitatLabs <- c(P = "Pier", SS = "Sediment")
# top_lol$Habitat[top_lol$Habitat == "P"] <-"Pier"
# top_lol$Habitat[top_lol$Habitat == "SS"] <-"Sediment"

p_top_lol <- top_lol %>% 
  ungroup() %>%
  mutate(Label=as.factor(Label),
  Label=fct_reorder(Label, mean)) %>%
  dplyr::select(-Pier, -Sediment) %>%
  ggplot( aes(x=sum, y=Label)) +
  scale_x_log10() +
  geom_linerange( aes(xmin=0, xmax=sum, y=Label, colour=Habitat), position=position_dodge(.5)) +
  geom_point( aes(colour=Habitat), position=position_dodge(.5), size=2) +
   # labs(title="Most abundant debris items",
   #     subtitle="Total counts by habitat (log10)",
   #     x="Total count", y="Debris item types", tag="C") +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        panel.background = element_blank()) 


p_top_lol

# ggsave(filename = "images/p_toplol.svg", plot = p_top_lol, device="svg")

svg("images/p_toplol.svg")
plot(p_top_lol)
dev.off()
```

## GLMM negative binomial model
I aggregated debris totals by survey date and location and used a negative binomial GLMM model to fit the data.  Debris totals are predicted by habitat, distance from harbour mouth, and aspect (north or south coast).  Location and Date are random effects.  

```{r glmm-hab, warning=FALSE, message=FALSE}

#create df to use for models
mod_hab <- site_select %>%
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) %>%
  ungroup() %>%
  mutate(Location = as.factor(Location),
         Habitat = as.factor(Habitat))


m_hab  <- glmmTMB(Total ~ #Debris count
                 Habitat * DistKm + #predicted by Habitat
                 (1|Location) + (1|Date), #date random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_hab)

drop1_hab <- drop1(m_hab, test="Chisq")
drop1_hab

# Predictors are: Habitat + DistKm

nd_hab <- expand.grid(Habitat = unique(mod_hab$Habitat),
                  DistKm = seq(from = min(mod_hab$DistKm),
                               to = max(mod_hab$DistKm),
                               length=1000),
                  Location = NA,
                  Date = NA)

pred_hab <- predict(object = m_hab,
                         newdata = nd_hab,
                         se.fit = T,
                         re.form = NA,
                         type="response")
                    

nd_hab$Total <- pred_hab$fit
nd_hab$SE_upper <- pred_hab$fit + pred_hab$se.fit
nd_hab$SE_lower <- pred_hab$fit - pred_hab$se.fit

#plot the predicted means
p_hab <- ggplot(nd_hab, aes(y=Total, x=DistKm)) + 
  geom_line(aes(col=Habitat)) +
  labs(title="Debris abundance predictions",

              subtitle="By habitat and distance from harbour mouth",
       x="Distance from harbour mouth (km)", y="Total debris items", tag="") +
  geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) 
    # theme(legend.position = "none",
    #     panel.grid = element_blank(),
    #     axis.title = element_blank(),
    #     axis.text = element_blank(),
    #     panel.background = element_blank()) 

p_hab

# ggsave(filename = "images/p_hab.svg", plot = p_hab, device="svg", height=7, width=12, units="in")

svg("images/p_hab.svg")
plot(p_hab, height=7, width=12, units="in")
dev.off()

citation("stats")

drop1(m_hab, test="Chisq") #interaction p=0.005621 **
```
```{r glmm-items}
options(scipen = 999)

# fishing line
      mod_line <- site_select %>%
        filter(Label=="Fishing line") %>% #fishing line
        group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
        summarise(Total=sum(Total)) %>%
        ungroup() %>%
                    mutate(Location = as.factor(Location),
               Habitat = as.factor(Habitat))
      
    #GLMM model and predictions
      m_line  <- glmmTMB(Total ~ #Debris count
                       Habitat + DistKm + #predicted by distance from harbour mouth
                       (1|Location) + (1|Date), #site random effect
                     family=nbinom1(), #negative binomial to deal with count data and zeros
                     data = mod_line) 
      
      nd_line <- expand.grid(Habitat = unique(mod_line$Habitat),
                        DistKm = seq(from = min(mod_line$DistKm),
                                     to = max(mod_line$DistKm),
                                     length=1000),
                        Location = NA,
                        Date = NA)
      
      pred_line <- predict(object = m_line,
                               newdata = nd_line,
                               se.fit = T,
                               re.form = NA,
                               type="response")
                          
      
      nd_line$Total <- pred_line$fit
      nd_line$SE_upper <- pred_line$fit + pred_line$se.fit
      nd_line$SE_lower <- pred_line$fit - pred_line$se.fit
      
    #plot the predicted means
      p_line <- ggplot(nd_line, aes(y=Total, x=DistKm)) + 
        geom_line(aes(col=Habitat)) +
        labs(title="Debris abundance predictions",
                    subtitle="By habitat and distance from harbour mouth",
             x="Distance from harbour mouth (km)", y="Total fishing line", tag="") +
        geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) 
      
      p_line
      
# Glass bottle
      mod_glbtl <- site_select %>%
        filter(Label=="Glass beverage bottle") %>% #glbtl
        group_by(Date, Location, Habitat) %>% #relevant variables
        summarise(Total=sum(Total)) %>%
        ungroup() %>%
                    mutate(Location = as.factor(Location),
               Habitat = as.factor(Habitat))
      
  #GLMM model and predictions
      m_glbtl  <- glmmTMB(Total ~ #Debris count
                       Habitat + #predicted by distance from harbour mouth
                       (1|Location) + (1|Date), #site random effect
                     family=nbinom1(), #negative binomial to deal with count data and zeros
                     data = mod_glbtl) 
      
      nd_glbtl <- expand.grid(Habitat = unique(mod_glbtl$Habitat),
                        Location = NA,
                        Date = NA)
      
      pred_glbtl <- predict(object = m_glbtl,
                               newdata = nd_glbtl,
                               se.fit = T,
                               re.form = NA,
                               type="response")
                          
      
      nd_glbtl$Total <- pred_glbtl$fit
      nd_glbtl$SE_upper <- pred_glbtl$fit + pred_glbtl$se.fit
      nd_glbtl$SE_lower <- pred_glbtl$fit - pred_glbtl$se.fit
      
  #plot the predicted means
      p_glbtl <- ggplot(nd_glbtl, aes(y=Total, x=Habitat)) + 
        geom_col(aes(fill=Habitat)) +
        labs(
          # title="Debris abundance predictions",
          #           subtitle="By habitat and distance from harbour mouth",
             x="Habitat", y="Total glass bottles", tag="") +
        geom_errorbar(aes(ymax = SE_upper, ymin = SE_lower), alpha=0.2) 
      
      p_glbtl
      
# Metal scraps
      mod_metsc <- site_select %>%
        filter(Label=="Metal scraps") %>% #metal scraps
        group_by(Date, Location, Habitat) %>% #relevant variables
        summarise(Total=sum(Total)) %>%
        ungroup() %>%
                    mutate(Location = as.factor(Location),
               Habitat = as.factor(Habitat))
      
  #GLMM model and predictions
      m_metsc  <- glmmTMB(Total ~ #Debris count
                       Habitat + #predicted by distance from harbour mouth
                       (1|Location) + (1|Date), #site random effect
                     family=nbinom1(), #negative binomial to deal with count data and zeros
                     data = mod_metsc) 
      
      nd_metsc <- expand.grid(Habitat = unique(mod_metsc$Habitat),
                        Location = NA,
                        Date = NA)
      
      pred_metsc <- predict(object = m_metsc,
                               newdata = nd_metsc,
                               se.fit = T,
                               re.form = NA,
                               type="response")
                          
      
      nd_metsc$Total <- pred_metsc$fit
      nd_metsc$SE_upper <- pred_metsc$fit + pred_metsc$se.fit
      nd_metsc$SE_lower <- pred_metsc$fit - pred_metsc$se.fit
      
  #plot the predicted means
      p_metsc <- ggplot(nd_metsc, aes(y=Total, x=Habitat)) + 
        geom_col(aes(fill=Habitat)) +
        labs(
          # title="Debris abundance predictions",
          #           subtitle="By habitat and distance from harbour mouth",
             x="Habitat", y="Total metal scraps", tag="") +
        geom_errorbar(aes(ymax = SE_upper, ymin = SE_lower), alpha=0.2) 
      
      p_metsc
      
# Fishhooks/sinkers
      mod_hook <- site_select %>%
        filter(Label=="Fishhook/sinker") %>% # fishhooks/sinkers
        group_by(Date, Location, Habitat) %>% #relevant variables
        summarise(Total=sum(Total)) %>%
        ungroup() %>%
                    mutate(Location = as.factor(Location),
               Habitat = as.factor(Habitat))
      
  #GLMM model and predictions
      m_hook  <- glmmTMB(Total ~ #Debris count
                       Habitat + #predicted by distance from harbour mouth
                       (1|Location) + (1|Date), #site random effect
                     family=nbinom1(), #negative binomial to deal with count data and zeros
                     data = mod_hook) 
      
      nd_hook <- expand.grid(Habitat = unique(mod_hook$Habitat),
                        Location = NA,
                        Date = NA)
      
      pred_hook <- predict(object = m_hook,
                               newdata = nd_hook,
                               se.fit = T,
                               re.form = NA,
                               type="response")
                          
      
      nd_hook$Total <- pred_hook$fit
      nd_hook$SE_upper <- pred_hook$fit + pred_hook$se.fit
      nd_hook$SE_lower <- pred_hook$fit - pred_hook$se.fit
      
  #plot the predicted means
      p_hook <- ggplot(nd_hook, aes(y=Total, x=Habitat)) + 
        geom_col(aes(fill=Habitat)) +
        labs(
          # title="Debris abundance predictions",
          #           subtitle="By habitat and distance from harbour mouth",
             x="Habitat", y="Total hooks & sinkers", tag="") +
        geom_errorbar(aes(ymax = SE_upper, ymin = SE_lower), alpha=0.2) 
      
      p_hook
      
# Fishing misc.
      mod_fmisc <- site_select %>%
        filter(Label=="Fishing misc.") %>% # fishfmiscs/sinkers
        group_by(Date, Location, Habitat) %>% #relevant variables
        summarise(Total=sum(Total)) %>%
        ungroup() %>%
                    mutate(Location = as.factor(Location),
               Habitat = as.factor(Habitat))
      
  #GLMM model and predictions
      m_fmisc  <- glmmTMB(Total ~ #Debris count
                       Habitat + #predicted by distance from harbour mouth
                       (1|Location) + (1|Date), #site random effect
                     family=nbinom1(), #negative binomial to deal with count data and zeros
                     data = mod_fmisc) 
      
      nd_fmisc <- expand.grid(Habitat = unique(mod_fmisc$Habitat),
                        Location = NA,
                        Date = NA)
      
      pred_fmisc <- predict(object = m_fmisc,
                               newdata = nd_fmisc,
                               se.fit = T,
                               re.form = NA,
                               type="response")
                          
      
      nd_fmisc$Total <- pred_fmisc$fit
      nd_fmisc$SE_upper <- pred_fmisc$fit + pred_fmisc$se.fit
      nd_fmisc$SE_lower <- pred_fmisc$fit - pred_fmisc$se.fit
      
  #plot the predicted means
      p_fmisc <- ggplot(nd_fmisc, aes(y=Total, x=Habitat)) + 
        geom_col(aes(fill=Habitat)) +
        labs(
          # title="Debris abundance predictions",
          #           subtitle="By habitat and distance from harbour mouth",
             x="Habitat", y="Total fmiscs & sinkers", tag="") +
        geom_errorbar(aes(ymax = SE_upper, ymin = SE_lower), alpha=0.2) 
      
      p_fmisc
```


```{r glmm-material}

# Metal predictions
      #create df of all metal
      mod_met <- site_select %>%
        filter(Material=="Metal") %>% #glass
        group_by(Date, Location, Habitat) %>% #relevant variables
        summarise(Total=sum(Total)) %>%
        ungroup() %>%
        mutate(Location = as.factor(Location),
               Habitat = as.factor(Habitat))
      
      
      m_met  <- glmmTMB(Total ~ #Debris count
                       Habitat + #predicted by Habitat
                       (1|Location) + (1|Date), #site/date random effect
                     family=nbinom1(), #negative binomial to deal with count data and zeros
                     data = mod_met)

      nd_met <- expand.grid(Habitat = unique(mod_met$Habitat),
                        Location = NA,
                        Date = NA)
      
      pred_met <- predict(object = m_met,
                               newdata = nd_met,
                               se.fit = T,
                               re.form = NA,
                               type="response")
                          
      
      nd_met$Total <- pred_met$fit
      nd_met$SE_upper <- pred_met$fit + pred_met$se.fit
      nd_met$SE_lower <- pred_met$fit - pred_met$se.fit
      
  #plot the predicted means
      p_met <- ggplot(nd_met, aes(y=Total, x=Habitat)) + 
        geom_col(aes(fill=Habitat)) +
        labs(
          # title="Debris abundance predictions",
          #           subtitle="By habitat and distance from harbour mouth",
             x="Habitat", y="Total metal items", tag="") +
        geom_errorbar(aes(ymax = SE_upper, ymin = SE_lower), alpha=0.2) 
      
      p_met
      
#Plastic predictions
      
      mod_plas <- site_select %>%
      filter(Material=="Plastic") %>% #glass
      group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
      summarise(Total=sum(Total)) %>%
      ungroup() %>%
      mutate(Location = as.factor(Location),
             Habitat = as.factor(Habitat))
    
    
      m_plas  <- glmmTMB(Total ~ #Debris count
                     Habitat*DistKm + #predicted by Habitat
                     (1|Location) + (1|Date), #site/date random effect
                   family=nbinom1(), #negative binomial to deal with count data and zeros
                   data = mod_plas)
      
      nd_plas <- expand.grid(Habitat = unique(mod_plas$Habitat),
                        DistKm = seq(from = min(mod_plas$DistKm),
                                     to = max(mod_plas$DistKm),
                                     length=1000),
                        Location = NA,
                        Date = NA)
      
      pred_plas <- predict(object = m_plas,
                               newdata = nd_plas,
                               se.fit = T,
                               re.form = NA,
                               type="response")
                          
      
      nd_plas$Total <- pred_plas$fit
      nd_plas$SE_upper <- pred_plas$fit + pred_plas$se.fit
      nd_plas$SE_lower <- pred_plas$fit - pred_plas$se.fit
      
    #plot the predicted means
      p_plas <- ggplot(nd_plas, aes(y=Total, x=DistKm)) + 
        geom_line(aes(col=Habitat)) +
        labs(title="Debris abundance predictions",
                    subtitle="By habitat and distance from harbour mouth",
             x="Distance from harbour mouth (km)", y="Total plastic items", tag="") +
        geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) 
      
      p_plas
      
#Glass predictions
            
      #create df of all glass
      mod_gls <- site_select %>%
        filter(Material=="Glass") %>% #glass
        group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
        summarise(Total=sum(Total)) %>%
        ungroup() %>%
        mutate(Location = as.factor(Location),
               Habitat = as.factor(Habitat))
      
      
      m_gls  <- glmmTMB(Total ~ #Debris count
                       Habitat*DistKm + #predicted by Habitat
                       (1|Location) + (1|Date), #site/date random effect
                     family=nbinom1(), #negative binomial to deal with count data and zeros
                     data = mod_gls)
      
      nd_gls <- expand.grid(Habitat = unique(mod_gls$Habitat),
                        DistKm = seq(from = min(mod_gls$DistKm),
                                     to = max(mod_gls$DistKm),
                                     length=1000),
                        Location = NA,
                        Date = NA)
      
      pred_gls <- predict(object = m_gls,
                               newdata = nd_gls,
                               se.fit = T,
                               re.form = NA,
                               type="response")
                          
      
      nd_gls$Total <- pred_gls$fit
      nd_gls$SE_upper <- pred_gls$fit + pred_gls$se.fit
      nd_gls$SE_lower <- pred_gls$fit - pred_gls$se.fit
      
    #plot the predicted means
      p_gls <- ggplot(nd_gls, aes(y=Total, x=DistKm)) + 
        geom_line(aes(col=Habitat)) +
        labs(title="Debris abundance predictions",
                    subtitle="By habitat and distance from harbour mouth",
             x="Distance from harbour mouth (km)", y="Total glass items", tag="") +
        geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) 
      
      p_gls
```
```{r print-mat-figs}
# run chunk 9 first

#glass material type
        p_gls <- ggplot(nd_gls, aes(y=Total, x=DistKm)) + 
                geom_line(aes(col=Habitat)) +
                labs(title="Debris abundance predictions",
                            subtitle="By habitat and distance from harbour mouth",
                            x="Distance from harbour mouth (km)", y="Total glass items", tag="") +
                geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) +
                theme(legend.position = "none",
                title = element_blank(),
                panel.grid = element_blank(),
                axis.title = element_blank(),
                axis.text = element_blank(),
                panel.background = element_blank())
        
        p_gls
        
        svg("images/p_gls.svg")
        plot(p_gls, height=7, width=7, units="in")
        dev.off()
        
#metal material type
        p_met <- ggplot(nd_met, aes(y=Total, x=Habitat)) + 
        geom_col(aes(fill=Habitat)) +
        labs(title="Debris abundance predictions",
             subtitle="By habitat and distance from harbour mouth",
             x="Habitat", y="Total metal items", tag="") +
        geom_errorbar(aes(ymax = SE_upper, ymin = SE_lower), alpha=0.2) +
        theme(legend.position = "none",
                title = element_blank(),
                panel.grid = element_blank(),
                axis.title = element_blank(),
                axis.text = element_blank(),
                panel.background = element_blank())
        
        p_met
        
        svg("images/p_met.svg")
        plot(p_met, height=7, width=7, units="in")
        dev.off()
        
#plastic material type
        p_plas <- ggplot(nd_plas, aes(y=Total, x=DistKm)) + 
        geom_line(aes(col=Habitat)) +
        labs(title="Debris abundance predictions",
                    subtitle="By habitat and distance from harbour mouth",
                    x="Distance from harbour mouth (km)", y="Total plastic items", tag="") +
        geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) +
        theme(legend.position = "none",
                title = element_blank(),
                panel.grid = element_blank(),
                axis.title = element_blank(),
                axis.text = element_blank(),
                panel.background = element_blank())
      
      p_plas
      
      svg("images/p_plas.svg")
        plot(p_plas, height=7, width=7, units="in")
        dev.off()
```
```{r glmm-fish, warning=FALSE, message=FALSE}

# Fishing items
      #create df to use for models
      mod_fish <- site_select %>%
        filter(Fishing=="Fishing") %>% #fishing line
        group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
        summarise(Total=sum(Total)) %>%
        ungroup() %>%
                mutate(Location = as.factor(Location),
               Habitat = as.factor(Habitat))
      
      #GLM negative binomial with two variables and zero inflation
      m_fish  <- glmmTMB(Total ~ #Debris count
                       Habitat * DistKm + #predicted by distance from harbour mouth
                       (1|Location) + (1|Date), #site random effect
                     family=nbinom1(), #negative binomial to deal with count data and zeros
                     data = mod_fish) 
      
      # Predictors for fishing related debris are: Habitat + DistKm
      
      nd_fish <- expand.grid(Habitat = unique(mod_fish$Habitat),
                        DistKm = seq(from = min(mod_fish$DistKm),
                                     to = max(mod_fish$DistKm),
                                     length=1000),
                        Location = NA,
                        Date = NA)
      
      pred_fish <- predict(object = m_fish,
                               newdata = nd_fish,
                               se.fit = T,
                               re.form = NA,
                               type="response")
      
      nd_fish$Total <- pred_fish$fit
      nd_fish$SE_upper <- pred_fish$fit + pred_fish$se.fit
      nd_fish$SE_lower <- pred_fish$fit - pred_fish$se.fit
      
      p_fish <- ggplot(nd_fish, aes(y=Total, x=DistKm)) + 
        geom_line(aes(col=Habitat)) +
        geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) 
        # theme(legend.position = "none",
        #       panel.grid = element_blank(),
        #       axis.title = element_blank(),
        #       axis.text = element_blank(),
        #       panel.background = element_blank()) 
      
      p_fish
      
      # svg("images/p_fish.svg")
      # plot(p_fish, height=7, width=7, units="in")
      # dev.off()


# Nonfishing items

      mod_nonfish <- site_select %>%
        filter(Fishing=="Non-Fishing") %>% #nonfishing debris
        group_by(Date, Location, Habitat) %>% #relevant variables
        summarise(Total=sum(Total)) %>%
        ungroup() %>%
                    mutate(Location = as.factor(Location),
               Habitat = as.factor(Habitat))
      
      m_nonfish  <- glmmTMB(Total ~ #Debris count
                       Habitat + #predicted by Habitat
                       (1|Location) + (1|Date), #site random effect
                     family=nbinom1(), #negative binomial to deal with count data and zeros
                     data = mod_nonfish)
      
      nd_nonfish <- expand.grid(Habitat = unique(mod_nonfish$Habitat),
                        Location = NA,
                        Date = NA)
      
      pred_nonfish <- predict(object = m_nonfish,
                               newdata = nd_nonfish,
                               se.fit = T,
                               re.form = NA,
                               type="response")
      
      nd_nonfish$Total <- pred_nonfish$fit
      nd_nonfish$SE_upper <- pred_nonfish$fit + pred_nonfish$se.fit
      nd_nonfish$SE_lower <- pred_nonfish$fit - pred_nonfish$se.fit
      
      #plot the predicted means
            p_nonfish <- ggplot(nd_nonfish, aes(y=Total, x=Habitat)) + 
            geom_col(aes(fill=Habitat)) +
            labs(
                # title="Debris abundance predictions",
                #           subtitle="By habitat and distance from harbour mouth",
                   x="Habitat", y="Total nonfishal items", tag="") +
            geom_errorbar(aes(ymax = SE_upper, ymin = SE_lower), alpha=0.2) 
            
            p_nonfish

```
```{r print-fish-figs}
# run chunk 11 glmm-fish first

#classed as fishing related
        p_fish <- ggplot(nd_fish, aes(y=Total, x=DistKm)) + 
                geom_line(aes(col=Habitat)) +
                labs(title="Debris abundance predictions",
                            subtitle="By habitat and distance from harbour mouth",
                            x="Distance from harbour mouth (km)", y="Total fishing items", tag="") +
                geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) +
                theme(legend.position = "none",
                title = element_blank(),
                panel.grid = element_blank(),
                axis.title = element_blank(),
                axis.text = element_blank(),
                panel.background = element_blank())
        
        p_fish
        
        svg("images/p_fish.svg")
        plot(p_fish, height=7, width=7, units="in")
        dev.off()
        
#classed as nonfishing related
        p_nonfish <- ggplot(nd_nonfish, aes(y=Total, x=Habitat)) + 
        geom_col(aes(fill=Habitat)) +
        labs(title="Debris abundance predictions",
             subtitle="By habitat and distance from harbour mouth",
             x="Habitat", y="Total nonfish items", tag="") +
          geom_errorbar(aes(ymax = SE_upper, ymin = SE_lower), alpha=0.2) +
        theme(legend.position = "none",
                title = element_blank(),
                panel.grid = element_blank(),
                axis.title = element_blank(),
                axis.text = element_blank(),
                panel.background = element_blank())
        
        p_nonfish
        
        svg("images/p_nonfish.svg")
        plot(p_nonfish, height=7, width=7, units="in")
        dev.off()
```

## Testing non-fishing related debris
```{r glm-nonfish, warning=FALSE, message=FALSE}
#create df to use for models
mod_nonfish <- site_select %>%
  filter(Fishing=="Non-Fishing") %>% #nonfishing debris
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) %>%
  ungroup() %>%
              mutate(Location = as.factor(Location),
         Habitat = as.factor(Habitat))

m_nonfish  <- glmmTMB(Total ~ #Debris count
                 Habitat * DistKm + #predicted by Habitat
                 (1|Location) + (1|Date), #site random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_nonfish)

vignette("basics", "emmeans")

library(emmeans)
emmeans(m_nonfish, pairwise~ Habitat|DistKm , at=list(DistKm=c(1,8)))

#pier diff 3.05 - 4.28 (-1.23) at 1km, 2.94 - 4.06 (-1.12) at 8km
#sedi diff 1.22 - 3.14 (-1.92) at 1km, 1.91 - 3.43 (-1.52) at 8km

# DistKm = 1:
#  contrast estimate    SE df t.ratio p.value
#  P - SS      1.486 0.540 29 2.751   0.0101 
# 
# DistKm = 8:
#  contrast estimate    SE df t.ratio p.value
#  P - SS      0.829 0.441 29 1.877   0.0705 

# Predictors for non-fishing related debris are: Habitat

nd_nonfish <- expand.grid(Habitat = unique(mod_nonfish$Habitat),
                          DistKm = seq(from = min(mod_nonfish$DistKm),
                               to = max(mod_nonfish$DistKm),
                               length=1000),
                  Location = NA,
                  Date = NA)


pred_nonfish <- predict(object = m_nonfish,
                         newdata = nd_nonfish,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_nonfish$Total <- pred_nonfish$fit
nd_nonfish$SE_upper <- pred_nonfish$fit + pred_nonfish$se.fit
nd_nonfish$SE_lower <- pred_nonfish$fit - pred_nonfish$se.fit


p_nonfish <- ggplot(nd_nonfish, aes(y=Total, x=DistKm)) + 
  geom_line(aes(col=Habitat)) +
  geom_ribbon(aes(ymin=SE_lower, ymax=SE_upper, fill=Habitat), alpha=0.2) +
   theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        panel.background = element_blank())

p_nonfish



# ggsave(filename = "images/p_nonfish.svg", plot = p_nonfish, device="svg", height=7, width=7, units="in")

svg("images/p_nonfish.svg")
plot(p_nonfish, height=7, width=7, units="in")
dev.off()

```

# glm plastic
```{r glm-plastic, warning=FALSE, message=FALSE}
#create df of all plastic materials
mod_plas <- site_select %>%
  filter(Material=="Plastic") %>% 
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) 

#GLM negative binomial with two variables and zero inflation
m_plas  <- glmmTMB(Total ~ #Debris count
                 Habitat*DistKm + #predicted by Habitat
                 (1|Location) + (1|Date), #site/date random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_plas) 

# Predictors for plastic are: Habitat, DistKm

nd_plas <- expand.grid(Habitat = unique(mod_plas$Habitat),
                          DistKm = seq(from = min(mod_plas$DistKm),
                               to = max(mod_plas$DistKm),
                               length=1000),
                  Location = NA,
                  Date = NA)

pred_plas <- predict(object = m_plas,
                         newdata = nd_plas,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_plas$Total <- pred_plas$fit
nd_plas$SE_upper <- pred_plas$fit + pred_plas$se.fit
nd_plas$SE_lower <- pred_plas$fit - pred_plas$se.fit

#plot the predicted means

p_plas <- ggplot(nd_plas, aes(y=Total, x=DistKm)) +
  geom_line(aes(col=Habitat)) +
  geom_ribbon(aes(ymax = SE_upper, ymin = SE_lower, fill=Habitat), alpha=0.2) 
   # theme(legend.position = "none",
   #      panel.grid = element_blank(),
   #      axis.title = element_blank(),
   #      axis.text = element_blank(),
   #      panel.background = element_blank()) 

p_plas

# ggsave(filename = "images/p_plas.svg", plot = p_plas, device="svg", height=7, width=7, units="in")

# svg("images/p_plas.svg")
# plot(p_plas, height=7, width=7, units="in")
# dev.off()

```
# glm glass
```{r glm-glsmat, warning=FALSE, message=FALSE}
#create df of all misc materials
mod_gls <- site_select %>%
  filter(Material=="Glass") %>% #glass
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) 


m_gls  <- glmmTMB(Total ~ #Debris count
                 Habitat*DistKm + #predicted by Habitat
                 (1|Location) + (1|Date), #site/date random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_gls)

# Predictors for 'other' materials are: Habitat, Coast

nd_gls <- expand.grid(Habitat = unique(mod_gls$Habitat),
                      DistKm = seq(from = min(mod_gls$DistKm),
                               to = max(mod_gls$DistKm),
                               length=1000),
                  Location = NA,
                  Date = NA)

pred_gls <- predict(object = m_gls,
                         newdata = nd_gls,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_gls$Total <- pred_gls$fit
nd_gls$SE_upper <- pred_gls$fit + pred_gls$se.fit
nd_gls$SE_lower <- pred_gls$fit - pred_gls$se.fit

#plot the predicted means

p_gls <- ggplot(nd_gls, aes(y=Total, x=DistKm)) +
  geom_line(aes(col=Habitat)) +
  geom_ribbon(aes(ymin=SE_lower, ymax=SE_upper, fill=Habitat), alpha=0.2) 
   # theme(legend.position = "none",
   #      panel.grid = element_blank(),
   #      axis.title = element_blank(),
   #      axis.text = element_blank(),
   #      panel.background = element_blank())

p_gls

# ggsave(filename = "images/p_gls.svg", plot = p_gls, device="svg", height=7, width=7, units="in")

svg("images/p_gls.svg")
plot(p_gls, height=7, width=7, units="in")
dev.off()

```


# glm metal
```{r glm-metmat, warning=FALSE, message=FALSE}
#create df of all misc materials
mod_met <- site_select %>%
  filter(Material=="Metal") %>% #glass
  group_by(Date, Location, Habitat, DistKm) %>% #relevant variables
  summarise(Total=sum(Total)) 

# #GLM negative binomial with two variables and zero inflation
m_met  <- glmmTMB(Total ~ #Debris count
                 Habitat*DistKm + #predicted by Habitat
                 (1|Location) + (1|Date), #site/date random effect
               family=nbinom1(), #negative binomial to deal with count data and zeros
               data = mod_met)

# Predictors for 'other' materials are: Habitat, Coast

nd_met <- expand.grid(Habitat = unique(mod_met$Habitat),
                      DistKm = seq(from = min(mod_met$DistKm),
                               to = max(mod_met$DistKm),
                               length=1000),
                  Location = NA,
                  Date = NA)

pred_met <- predict(object = m_met,
                         newdata = nd_met,
                         se.fit = T,
                         re.form = NA,
                         type="response")

nd_met$Total <- pred_met$fit
nd_met$SE_upper <- pred_met$fit + pred_met$se.fit
nd_met$SE_lower <- pred_met$fit - pred_met$se.fit

#plot the predicted means

p_met <- ggplot(nd_met, aes(y=Total, x=DistKm)) +
  geom_line(aes(col=Habitat)) +
  geom_ribbon(aes(ymin=SE_lower, ymax=SE_upper, fill=Habitat), alpha=0.2) +
  ylim(0, 40)
   # theme(legend.position = "none",
   #      panel.grid = element_blank(),
   #      axis.title = element_blank(),
   #      axis.text = element_blank(),
   #      panel.background = element_blank())

p_met

svg("images/p_met.svg")
plot(p_met, height=7, width=7, units="in")
dev.off()

# ggsave(filename = "images/p_met.svg", plot = p_met, device="svg", height=7, width=7, units="in")
```
## Ordination
```{r vegan-20}

#Abundance data
  #Filter for top items
top20 <- site_select %>%
  group_by(Label) %>%
  summarise(sum=sum(Total)) %>%
  arrange(desc(sum)) %>%
  top_n(20, sum)

 #create vector from top 20 items
 top20 <- top20$Label
 top20 <- unlist(top20)

  #create abundance dataframe with ID as columns and survey events as rows
  abund_long <- site_select %>%
    mutate(survey_event=paste0(Date, Site)) %>% #create column for unique survey event
    group_by(survey_event, DistKm, Habitat, Date, Label) %>% 
    summarise(sum=sum(Total)) %>% #total of each item type per survey event
  group_by(Label) %>%
  filter(sum>0) %>% #filter IDs that have no values across all surveys
  ungroup()
  
  #convert to data wide
  abund <- spread(abund_long, Label, sum) 
  
  #replace na values with 0
  abund[is.na(abund)] <- 0
  
  #convert survey event to rowname
  length(unique(abund$survey_event)) == nrow(abund) #checking for duplicates
  abund <- column_to_rownames(abund, var="survey_event") #create rownames from survey event
  
  com <- abund[4:64]
  env <- abund[1:2]
  
  ord <- metaMDS(com)
  
  ord$stress 
  #0.1977957
  
  # https://jkzorz.github.io/2020/04/04/NMDS-extras.html
  en = envfit(ord, env, permutations = 999, na.rm = TRUE)
  en
  
en_coord_cont = as.data.frame(scores(en, "vectors")) * ordiArrowMul(en)
en_coord_cat = as.data.frame(scores(en, "factors")) * ordiArrowMul(en)
  
  data.scores <- as.data.frame(scores(ord)) #create df of ord points
  data.scores$DistKm <- abund$DistKm
  data.scores$Habitat <- abund$Habitat
  
  species.scores <- as.data.frame(scores(ord, "species"))
  species.scores$item <- rownames(species.scores) #create item id column


#create ellipsis
#add group column to data.scores
NMDS <- data.frame(NMDS1 = ord$points[,1], NMDS2 = ord$points[,2],group=data.scores$Habitat)
 veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
  {
    theta <- (0:npoints) * 2 * pi/npoints
    Circle <- cbind(cos(theta), sin(theta))
    t(center + scale * t(Circle %*% chol(cov)))
  }
  df_ell <- data.frame()
  for(g in levels(NMDS$group)){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$group==g,],
                    veganCovEllipse(cov.wt(cbind(NMDS1,NMDS2),wt=rep(1/length(NMDS1),length(NMDS1)))$cov,center=c(mean(NMDS1),mean(NMDS2)))))
                    ,group=g))
  }

  p_ord <- ggplot() +
  # geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2, label=item)) +
  geom_point(data=data.scores, 
             aes(x=NMDS1, y=NMDS2, colour=Habitat), alpha=0.8) + 
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), 
            size=1, linetype=2) +
    geom_segment(aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),
       data = en_coord_cont, size =1, alpha = 0.5, colour = "grey30") +
     geom_text(data = en_coord_cat, aes(x = NMDS1, y = NMDS2+0.04),
       label = row.names(en_coord_cat), colour = "navy", fontface = "bold") +
     geom_text(data = en_coord_cont, aes(x = NMDS1, y = NMDS2), colour = "grey30",
       fontface = "bold", label = row.names(en_coord_cont)) +
  coord_equal() +
   theme(legend.position = "none",
        panel.background = element_blank())
  
  p_ord

svg("images/p_ord.svg", height=7, width=11)
plot(p_ord, height=7, width=11, units="in")
dev.off()

 txt20 <- species.scores %>%
   filter(item %in% top20)

  p_ord_txt <- ggplot() +
  geom_text(data=txt20,aes(x=NMDS1,y=NMDS2, label=item)) +
  geom_point(data=data.scores, 
             aes(x=NMDS1, y=NMDS2, colour=Habitat), alpha=0.8) + 
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=group), 
            size=1, linetype=2) +
  coord_equal() +
   theme(legend.position = "none",
        panel.background = element_blank())
  
  p_ord_txt

# svg("images/p_ord_txt.svg", height=7, width=11)
# plot(p_ord_txt, height=7, width=11, units="in")
# dev.off()

```

```{r adonis-pairwise, warning=FALSE, message=FALSE}
# https://jkzorz.github.io/2020/04/04/NMDS-extras.html

#Both hab and dist are significant, as is the interaction
df.adonis <- adonis(com ~ data.scores$Habitat*data.scores$DistKm, permutations=999, method="bray")
df.adonis

# https://search.r-project.org/CRAN/refmans/vegan/html/permutest.betadisper.html

#testing habitat dispersion
df.dis <- vegdist(com, method = "bray") # make a distance matrix
df.betadisper.hab <- betadisper(df.dis, data.scores$Habitat) # test dispersion
df.betadisper.hab
perm.hab <- permutest(df.betadisper.hab) 
#testing dist dispersion
df.dis <- vegdist(com, method = "bray") # make a distance matrix
df.betadisper.dist <- betadisper(df.dis, data.scores$DistKm) # test dispersion
df.betadisper.dist
perm.dist <- permutest(df.betadisper.dist) 

## Tukey's Honest Significant Differences
TKhab <- TukeyHSD(df.betadisper.hab)
plot(TKhab)

TKdist <- TukeyHSD(df.betadisper.dist)
plot(TKdist)


pstat.hab <- permustats(perm.hab)
densityplot(pstat.hab, scales = list(x = list(relation = "free")))
qqmath(pstat.hab, scales = list(relation = "free"))

pstat.dist <- permustats(perm.dist)
densityplot(pstat.dist, scales = list(x = list(relation = "free")))
qqmath(pstat.dist, scales = list(relation = "free"))

```
## gllvm analysis
Determining the significance of environmental traits and debris attributes.

Coefficient plot shows estimated coefficients (ticks) and their 95% confidence interval (lines). that there are a few significant variables including Habitat affected by fishing, habitat on its own, and the dispersiveness category of unencumbered drift.  

Here we see that nothing is interacting significantly with distance to harbour mouth, however both fishing and encumbered drift are interacting with habitat significantly.

```{r gllvm-set-up-1or0, warning=FALSE, message=FALSE, results="hide"}
#Abundance
  #create abundance dataframe with ID as columns and survey events as rows
  abund_long <- site_select %>%
    mutate(survey_event=paste0(Date, Site)) %>% #create column for unique survey event
    group_by(survey_event, ID) %>% 
    summarise(sum=sum(Total)) %>% #total of each item type per survey event
    dplyr::select(survey_event, ID, sum) %>%
  group_by(ID) %>%
  filter(sum>0) %>% #filter IDs that have no values across all surveys
  ungroup()
  
  #convert to data wide
  abund <- spread(abund_long, ID, sum) 
  
  #replace na values with 0
  abund[is.na(abund)] <- 0
  
  #convert survey event to rowname
  length(unique(abund$survey_event)) == nrow(abund) #checking for duplicates
  abund <- column_to_rownames(abund, var="survey_event") #create rownames from survey event

#Environment
  #create survey event and filter zero sum surveys, 
    env <- site_select %>%
    mutate(survey_event=paste0(Date, Site)) %>%
      group_by(survey_event) %>%
      mutate(sum=sum(Total)) %>%
      filter(sum>0) %>%
      dplyr::select(survey_event, Habitat, DistKm, Location) %>%
      distinct() %>%
      ungroup() %>%
      mutate(Pier = ifelse(Habitat=="P", 1, 0))
  
    
  #convert survey event to rowname
  length(unique(env$survey_event)) == nrow(env) #checking for duplicates
  env <- column_to_rownames(env, var="survey_event") #create rownames from survey event
  #convert Hab to numeric
  env$Pier <- as.numeric(env$Pier) #convert factor to numeric

  env <- env %>%
    dplyr::select(DistKm, Pier)

#Traits
  
  #create column of usable IDs to join
  IDs <- abund %>%
    gather(ID, sum) %>%
    dplyr::select(ID) %>%
    unique()
  #create trait df with ID type as row and columns are fishing and dispersiveness
  TR <- site_select %>%
    group_by(ID) %>%
    summarise(sum=sum(Total)) %>%
    left_join(item_code, by="ID") %>%
    dplyr::select(Material, ID, Dispersiveness, Fishing) %>%
    inner_join(IDs, by="ID") %>% #remove items with 0 count
    column_to_rownames(var="ID") #create rownames from ID
  
  #state variables to rename
  var_change <- c("Ceramic", "Undetermined", "E-Waste", "Paper", "Construction", "Organic", "Unknown", "Timber", "Foam", "Brick or Cement", "Rubber", "Cloth", "") 
  
  #change Material to character
  TR$Material <- as.character(TR$Material) 
  
  #set lesser used categories to 'Other'
  TR$Material[TR$Material %in% var_change] = "Other" 

  TR$Dispersiveness <- as.factor(TR$Dispersiveness) #convert character to factor with 4 levels
  TR$Fishing <- as.factor(TR$Fishing) #convert character to factor with 2 levels
  TR$Material <- as.factor(TR$Material) #convert character to factor with 4 levels
  
  #alphabatising variables to force intercepts
  TR$Material <- recode_factor(TR$Material, Other = "AOther")
  TR$Fishing <- recode_factor(TR$Fishing, 'Non-fishing' = "ANon-fishing")
  TR$Dispersiveness <- recode_factor(TR$Dispersiveness, 'Very mobile' = "AVery mobile")

y <- as.matrix(abund)
X <- scale(as.matrix(env))
TR <- TR

fit4th <- gllvm(y, X, TR, family="negative.binomial",
                formula = y ~ (Pier + DistKm) +
                  (Pier + DistKm) : (Dispersiveness + Fishing + Material))

gllvm(y, X, TR, family="negative.binomial")


coefplot(fit4th, cex.ylab = 0.8)

summary(fit4th)

fourth <- fit4th$fourth.corner
a <- max( abs(fourth))
colort <- colorRampPalette(c("blue", "white", "red"))
plot.4th <- levelplot((as.matrix(fourth)), 
                      xlab = "Environmental Variables", 
                      ylab = "Debris traits",
                      col.regions = colort(100), #use defined colour palette
                      cex.lab = 1.3, #specify size of the label text
                      at = seq(-a, a, length = 100), #makes heat scale equal both sides
                      scales = list(x = list(rot = 45))) #rotate text for clarity


plot.4th

summary(fourth)

png(file = "images/coefplot.png")  # The directory you want to save the file in
coefplot(fit4th, cex.ylab = .6)
dev.off()

#call svg command to start the plot
svg(file = "images/plot4th.svg",   # The directory you want to save the file in
    width = 11, # The width of the plot in inches
    height = 7) # The height of the plot in inches
plot(plot.4th)
dev.off()
```


```{r fixed-intercept}
fit4th_z <- gllvm(y, X, TR, family="negative.binomial",
                formula= y~ 0 + (Pier + DistKm) +
                  (Pier + DistKm) : (Dispersiveness + Fishing + Material))

summary(fit4th_z)

fourth_z <- fit4th_z$fourth.corner
a_z <- max( abs(fourth_z))
colort <- colorRampPalette(c("blue", "white", "red"))
plot.4th_z <- levelplot((as.matrix(fourth_z)), 
                      xlab = "Environmental Variables", 
                      ylab = "Debris traits",
                      col.regions = colort(100), #use defined colour palette
                      cex.lab = 1.3, #specify size of the label text
                      at = seq(-a, a, length = 100), #makes heat scale equal both sides
                      scales = list(x = list(rot = 45))) #rotate text for clarity


plot.4th_z

summary(fourth_z)
```

